In modern sports analytics, precise kinematic metrics such as ball speed, entry angle, trajectory curvature, and spin rate are fundamental for evaluating player performance. In the sport of bowling, extracting these metrics is particularly critical, as the physical interaction between the ball's rotational dynamics and the localized oil pattern of the lane dictates the ultimate success of a shot \cite{frohlich2004what}. 

Despite the importance of this data, extracting intricate parameters from live gameplay presents a significant technical barrier. Currently, commercial systems capable of extracting these metrics typically rely on complex, synchronized multi-camera stereoscopic setups or specialized sensor arrays \cite{gowda2017bringing}. While highly accurate, these proprietary systems are cost-prohibitive, restricting access strictly to professional broadcasting environments or high-end training facilities. Consequently, there is a growing demand to democratize sports analytics by developing robust computer vision systems capable of extracting deep metric insights from standard, accessible camera hardware. 

However, transitioning from specialized sensor arrays to a single, uncalibrated static camera introduces a host of geometric and visual challenges. Because intrinsic camera parameters are typically unavailable in broadcast or consumer footage, resolving depth requires strict adherence to projective geometry and physical constraints. Furthermore, tracking fast-moving sports balls from a monocular feed presents multiple challenges due to motion blur, occlusion, and background clutter, making the calculation of sub-pixel rotational velocity (spin) highly complex.

To bridge this gap, this report proposes a comprehensive, monocular computer vision pipeline designed to extract robust 3D trajectory and spin parameters from standard bowling video footage. 
